Minimax-optimal Laplacian Eigenmaps regression over Sobolev Spaces with Neighborhood Graphs

### 2021-03-21 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Finished editing the section on rates for manifold domain.

Broadly speaking, we show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2$-norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing. This statement holds for different relations between the dimension $d$ and number of derivatives $s$, depending on the problem (estimation or testing).
\begin{table}
	\begin{center}
		\begin{tabular}{p{.2\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Smoothness order & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
			\hline
			$s \leq 3$ & $\bm{n^{-2s/(2s + d)}}$ & $\bm{n^{-2s/(2s + m)}}$ \\
			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & \textcolor{red}{(?)}
		\end{tabular}
	\end{center}
	\caption{Summary of estimation rates over Sobolev balls. Bold font marks minimax optimal rates. In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $d, m \in \mathbb{N}$ (under Model~\ref{def:model_manifold}). Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
	\label{tbl:estimation_rates}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
			\hline
			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & \textcolor{red}{(?)} \\
			& $4 < \dim(\mc{X}) < 4s$ & \textcolor{red}{(?)} & \textcolor{red}{(?)} \\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
		\end{tabular}
	\end{center}
	\caption{Summary of testing rates over Sobolev balls. Bold font marks minimax optimal rates. Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$. Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
	\label{tbl:testing_rates}
\end{table}

@@

When $4s \geq d$ the functions in $H^s(\Xset)$ are very irregular---formally speaking $H^s(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $4s \geq d$---and as far as we know, the minimax testing rates in this regime have not been worked out. However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$ then the critical radius is given by $\epsilon^2 \asymp Mn^{-1/2}$; as a sanity check, note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}. As we discuss after our first main theorem regarding testing (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.

@@

As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral series method, of particular use when the design distribution $P$ is unknown. The hope is that the error incurred by this approximation is sufficiently small such that Laplacian eigenmaps preserves the optimality properties of the spectral series methods they are approximating. In this section and the following one, we will show that this is indeed the case; the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.

@@

Some other remarks:
\begin{itemize}
	\item When $M \ll n^{-1/2}$ or $M \gg n^{1/d}$, the rate in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} looks different than~\eqref{eqn:sobolev_space_minimax_estimation_rate}. In particular, when $M \ll n^{-1/2}$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$, but neither match the minimax rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}. On the other hand, when $M \gg n^{1/d}$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \asymp 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}. However, in truth these are edge cases, which do not fall neatly into the framework of nonparametric regression.
	\item The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected. On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is on the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$. Finally, we choose $K \asymp (M^2n)^{2d/(2 + d)}$ to optimally trade-off bias and variance.
	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$. In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation. However, because the estimator $\wh{f}$ is defined only in-sample, we cannot use such methods to select the graph radius $\varepsilon$, or number of eigenvectors $K$. We return to this issue upon considering out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
	\item The upper bound given in Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$. Under the stronger assumption that $f_0 \in C^1(\mc{X};M)$ we can replace the factor of $\delta$ by the sharper $\delta^2/n$, which is less than $\delta$ because $\delta \in (0,1)$. 
\end{itemize}

@@

\subsection{Higher-order Sobolev classes}
\label{sec:higher_order_sobolev_classes}
We now assume that the regression function displays some higher-order regularity and is zero-trace, i.e. $f_0 \in H_0^s(P)$. At first glance, it is not at all clear that Laplacian eigenmaps \emph{can} take advantage of this additional regularity, and it is especially unclear whether it can do so in an optimal manner. As $s \to \infty$ and $d$ remains fixed, the optimal estimation and testing rates approach the parametric rate $n^{-1}$. On the other hand, the sharpest known results show that Laplacian eigenvectors $v_k$ converge to continuum eigenfunctions $\psi_k$ only at a rate of \textcolor{red}{(?)}, which is obviously much slower than $n^{-1}$.

@@

	\item \textcolor{red}{(TODO 1): This entire comment is too vague right now. I'll leave it in, in case Siva + Ryan think it is sufficiently valuable to edit and make more precise.} The lower bound on $\varepsilon$ means that the graph $G$ will be quite dense, with average degree growing polynomially in $n$ as $n \to \infty$. Consequently, exactly computing the eigenvectors $v_k$ is more computationally intensive than if the graph radius were on the order of the connectivity threshold $\varepsilon \asymp (\log n/n)^{1/d}$. Recall that in the first-order case, we established in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that this smaller choice of $\varepsilon$ still resulted in an optimal estimator. Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be viewed as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds. 

@@

	Suppose one does choose $\varepsilon$ meaningfully larger than the connectivity threshold, as our theory requires when $s > 1$. We now discuss a procedure to efficiently compute an approximation to the Laplacian eigenmaps estimate, without meaningfully increasing the statistical error incurred: \emph{edge sparsification}. For a given $\delta \in (0,1)$, various algorithms have been designed \textcolor{red}{(references)} to efficiently remove many edges from the graph $G$, resulting in a sparser graph $\wc{G}$ with Laplacian $\wc{L}_{n,\varepsilon}$, while only perturbing the spectrum of the Laplacian by a factor of at most $\sigma \geq 1$, meaning
	\begin{equation*}
	\frac{1}{\sigma} \cdot u^{\top} \wc{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wc{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
	\end{equation*}
	The graph $\wc{G}$ is said to be $\sigma$-spectrally similar to $G$. Let $\wc{f}$ be the Laplacian eigenmaps estimator, computed with respect to the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ . Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$ \textcolor{red}{(reference needed)}. Statistically speaking, as we show in Appendix~\textcolor{red}{(?)} $\wc{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times the error of $\wh{f}$. Therefore, as long as $\sigma \asymp 1$ the estimator $\wc{f}$ will also be rate-optimal. Indeed, as pointed out by~\citet{sadhanala16b}, geometric graphs such as $G$ exhibit a special structure. In this case there exist far simpler and faster methods for sparsification than those of~\textcolor{red}{(references)}, which, at least empirically, appear to suffice.
	\item The requirement $p \in C^{s - 1}(\mc{X})$ is a strong condition, but is essential to showing that $\wh{f}$ enjoys faster rates of convergence. We give more detail about this when we discuss our analysis in Section~\ref{subsec:analysis}. 
\end{itemize}

\paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{3}
	\item 
	\label{asmp:parameters_testing_ho}
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_ho}. Additionally,
	\begin{equation}
	\label{eqn:eigenvector_testing_ho}
	K \asymp \min\Bigl\{(M^2n)^{2d/(4s + d)} \vee 1, n\Bigr\}.
	\end{equation}
\end{enumerate}
When $d \leq 4$, for any value of $s$ it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that $\varphi$ is an optimal test. In what follows, let $p(d) = \min\{1,2s/d\}$. 
\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_ho}
	Fix $a,b \in (0,1)$. Suppose Model~\ref{def:model_flat_euclidean}. Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test. Suppose additionally $f_0 \in H_0^s(P,M)$,that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$. Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho}
	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n},
	\end{equation}
	then $\Ebb_{f_0}[1 - \phi] \leq b$.
\end{theorem}
Just as in the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(P)$ when $n^{-1/2} \ll M^2 \ll n^{1/4}$ (if $2s < d$), or if $n^{-1/2} \ll M^2 \ll (2/d - 1/2)$ (if $2s \geq d$.)

However, unlike the first-order case, when $4 < d < 4s$ the optimal testing rate over $H_0^s(P)$ continues to be $n^{-4s/(4s + d)}$. Unfortunately, we cannot claim that $\varphi$ is an optimal test in this regime. The culprit is that there no longer exist any choices of $\varepsilon,K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}, and we can therefore no longer balance testing bias and variance (defined momentarily in~\textcolor{red}{(?)}). From Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, we can gather that $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$. In fact, we can derive slightly better (though still not optimal) rates.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_ho_suboptimal}
	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $d > 4$. If the Laplacian eigenmaps test $\varphi$ is computed with $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon \asymp K^{-1/d}$, and if 
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}
	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-2s/(2(s - 1) + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n},
	\end{equation}
	then $\Ebb_{f_0}[1 - \phi] \leq b$.
\end{theorem}
We do not know whether the culprit here is the test itself, our some looseness in our theoretical bounds, and leave the matter for future work.

@@

The optimal rates for nonparametric regression suffer from a bona fide curse of dimensionality, and are thus difficult to reconcile with the practical success of many methods for nonparametric regression. One explanation that resolves this seeming paradox is the \emph{manifold hypothesis}, where the design points $X$ are assumed to lie on or near a manifold $\mc{X}$ of intrinsic dimension $m < d$. Under the manifold hypothesis nonparametric regression is demonstrably easier, and in particular it is known \citep{bickel2007,ariascastro2018} that in this setting the optimal rates over $H^s(\Xset)$ scale like $n^{-2s/(2s + m)}$ (for estimation) and $n^{-4s/(4s + m)}$ (for testing).

On the other hand, a theory has been developed~\citep{belkin03,belkin05,niyogi2013} establishing the the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear. In this section, we contribute to this line of work by showing that under the manifold hypothesis, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates. First, we give some relevant background on Riemmanian manifolds, and Sobolev classes defined upon them.

>>>>

Broadly speaking, we show that when the regression function $f_0$ is smooth, in the Sobolev sense of having weak derivatives bounded in $\Leb^2$-norm, Laplacian eigenmaps methods are statistically minimax optimal, for both estimation and goodness-of-fit testing. This statement holds for different relations between the dimension $d$ and number of derivatives $s$, depending on the problem (estimation or testing).
\begin{table}
	\begin{center}
		\begin{tabular}{p{.2\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Smoothness order & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
			\hline
			$s \leq 3$ & $\bm{n^{-2s/(2s + d)}}$ & $\bm{n^{-2s/(2s + m)}}$ \\
			$s > 3$  & $\bm{n^{-2s/(2s + d)}}$ & $n^{-6/(6 + m)}$
		\end{tabular}
	\end{center}
	\caption{Summary of estimation rates over Sobolev balls. Bold font marks minimax optimal rates. In each case, rates hold for all $d \in \mathbb{N}$ (under Model~\ref{def:model_flat_euclidean}), and for all $d, m \in \mathbb{N}$ (under Model~\ref{def:model_manifold}). Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
	\label{tbl:estimation_rates}
\end{table}

\begin{table}
	\begin{center}
		\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
			\hline
			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & $n^{-12/(12 + d)}$ \\
			& $4 < \dim(\mc{X}) < 4s$ & $n^{-2s/(2(s - 1) + d)}$ & $n^{-6/(4 + m)}$ \\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
		\end{tabular}
	\end{center}
	\caption{Summary of testing rates over Sobolev balls. Bold font marks minimax optimal rates. Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$. Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
	\label{tbl:testing_rates}
\end{table}

@@

When $4s \geq d$ the functions in $H^s(\Xset)$ are very irregular---formally speaking $H^s(\Xset)$ does not continuously embed into $\Leb^4(\Xset)$ when $4s \geq d$---and as far as we know, the minimax testing rates in this regime have not been worked out. However, if one explicitly assumes that $f_0 \in \Leb^4(\mc{X})$ then the critical radius is given by $\epsilon^2 \asymp Mn^{-1/2}$. As a sanity check note that this is strictly worse than the rate in~\eqref{eqn:sobolev_space_testing_critical_radius}. As we discuss after our first main theorem regarding testing (Theorem~\ref{thm:laplacian_eigenmaps_testing_fo}), this rate is achievable by a test based on $\wh{T}$.

@@

As previously explained, Laplacian eigenmaps is a discrete and noisy approximation to a spectral series method, of particular use when the design distribution $P$ is unknown. Our goal is to show that Laplacian eigenmaps are rate optimal methods, notwithstanding the extra error incurred by this approximation. In this section and the following one, we will see that this is indeed the case: the estimator $\wh{f}$, and a test using the statistic $\wh{T}$, achieve optimal estimation and goodness-of-fit testing rates over Sobolev classes.

@@

Some other remarks:
\begin{itemize}
	\item When $M \ll n^{-1/2}$ or $M \gg n^{1/d}$, the rate in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} looks different than~\eqref{eqn:sobolev_space_minimax_estimation_rate}. In particular, when $M \ll n^{-1/2}$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$, but neither match the minimax rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}. On the other hand, when $M \gg n^{1/d}$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \asymp 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}. However, in truth these are edge cases, which do not fall neatly into the framework of nonparametric regression.
	\item The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected. On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is on the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$. Finally, we choose $K \asymp (M^2n)^{2d/(2 + d)}$ to optimally trade-off bias and variance.
	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$. In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation. However, because the estimator $\wh{f}$ is defined only in-sample, we cannot use such methods to select the graph radius $\varepsilon$, or number of eigenvectors $K$. We return to this issue in Section~\ref{sec:out_of_sample}, when we propose an out-of-sample extension of $\wh{f}$.
	\item The upper bound given in Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$. Under the stronger assumption that $f_0 \in C^1(\mc{X};M)$ we can replace the factor of $\delta$ by the sharper $\delta^2/n$, which is less than $\delta$ because $\delta \in (0,1)$. 
\end{itemize}

@@

\subsection{Higher-order Sobolev classes}
\label{sec:higher_order_sobolev_classes}
We now assume that the regression function displays some higher-order regularity and is zero-trace, i.e. $f_0 \in H_0^s(P)$. At first glance, it is not at all clear that Laplacian eigenmaps \emph{can} take advantage of this additional regularity, and it is especially unclear whether it can do so in an optimal manner. As $s \to \infty$ and $d$ remains fixed, the optimal estimation and testing rates approach the parametric rate $n^{-1}$. On the other hand, the sharpest known results show that Laplacian eigenvectors $v_k$ converge to continuum eigenfunctions $\psi_k$---and thus, for a fixed value of $k$, $\wh{f}$ converges to $\wt{f}$---only at a rate of \textcolor{red}{(?)}, which is obviously much slower than $n^{-1}$.

@@

	\item \textcolor{red}{(TODO 1): This entire comment is too vague right now. I'll leave it in, in case Siva + Ryan think it is sufficiently valuable to edit and make more precise.} The lower bound on $\varepsilon$ means that the graph $G$ will be quite dense, with average degree growing polynomially in $n$ as $n \to \infty$. Consequently, exactly computing the eigenvectors $v_k$ is more computationally intensive than if the graph radius were on the order of the connectivity threshold $\varepsilon \asymp (\log n/n)^{1/d}$. Recall that in the first-order case, we established in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that as long as $\varepsilon \gg (\log n/n)^{1/d}$---meaning average degree growing like $\omega(\log n)$---the estimator $\wh{f}$ could still be optimal. Thus Theorems~\ref{thm:laplacian_eigenmaps_estimation_fo} and~\ref{thm:laplacian_eigenmaps_estimation_ho} can be viewed as revealing a tradeoff between statistical and computational efficiency; although to be clear, we have no theoretical evidence that Laplacian eigenmaps \emph{fails} to adapt to higher-order smoothness when $\varepsilon \asymp (\log n/n)^{1/d}$---we simply cannot prove that it succeeds. 

@@

	Suppose one does choose $\varepsilon$ meaningfully larger than the connectivity threshold, as our theory requires when $s > 1$. We now discuss a procedure to efficiently compute an approximation to the Laplacian eigenmaps estimate, without meaningfully increasing the statistical error incurred: \emph{edge sparsification}. For a given $\delta \in (0,1)$, various algorithms have been designed \textcolor{red}{(references)} to efficiently remove many edges from the graph $G$, resulting in a sparser graph $\wc{G}$ with Laplacian $\wc{L}_{n,\varepsilon}$, while only perturbing the spectrum of the Laplacian by a factor of at most $\sigma \geq 1$, meaning
	\begin{equation*}
	\frac{1}{\sigma} \cdot u^{\top} \wc{L}_{n,\varepsilon} u \leq u^{\top} L_{n,\varepsilon} u \leq \sigma \cdot u^{\top} \wc{L}_{n,\varepsilon}u \quad \textrm{for all $u \in \Reals^n$.}
	\end{equation*}
	The graph $\wc{G}$ is said to be $\sigma$-spectrally similar to $G$. Let $\wc{f}$ be the Laplacian eigenmaps estimator computed using the eigenvectors of the sparsified graph Laplacian $\wc{L}_{n,\varepsilon}$ . Because $\wc{G}$ is sparser than $G$, it can be (much) faster to compute the eigenvectors of $\wc{L}_{n,\varepsilon}$ than the eigenvectors of $L_{n,\varepsilon}$, and consequently much faster to compute $\wc{f}$ than $\wh{f}$\textcolor{red}{(reference needed)}. Statistically speaking, as we show in Appendix~\textcolor{red}{(?)} $\wc{f}$ has $L^2(P_n)$-error of at most $\sigma^{2s}$ times the error of $\wh{f}$. Therefore, as long as $\sigma \asymp 1$ the estimator $\wc{f}$ will also be rate-optimal. Finally, geometric graphs such as $G$ exhibit a special structure; very roughly speaking, no one edge is a bottleneck in geometric graphs. As pointed out by~\citet{sadhanala16b}, in this case there exist far simpler and faster methods for sparsification than those of~\textcolor{red}{(references)}; at least empirically, these methods appear to do the job.
	\item The requirement $p \in C^{s - 1}(\mc{X})$ is a strong condition, but is essential to showing that $\wh{f}$ enjoys faster rates of convergence when $s > 1$. We give more detail about this when we discuss our analysis in Section~\ref{subsec:analysis}. 
\end{itemize}

\paragraph{Testing.} The test $\varphi$ can adapt to the higher-order smoothness of $f_0$, when $\varepsilon$ and $K$ are chosen correctly.
\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{3}
	\item 
	\label{asmp:parameters_testing_ho}
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_ho}. Additionally,
	\begin{equation}
	\label{eqn:eigenvector_testing_ho}
	K \asymp \min\Bigl\{(M^2n)^{2d/(4s + d)} \vee 1, n\Bigr\}.
	\end{equation}
\end{enumerate}
When $d \leq 4$, for any value of $s \in \mathbb{N}$ it is possible to choose $\varepsilon$ and $K$ such that both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho} are satisfied, and our next theorem establishes that in this situation $\varphi$ is an optimal test. In what follows, let $p(d) = \min\{1,2s/d\}$. 
\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_ho}
	Fix $a,b \in (0,1)$. Suppose Model~\ref{def:model_flat_euclidean}. Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test. Suppose additionally $f_0 \in H_0^s(P,M)$, that $p \in C^{s-1}(\mc{X})$, and that $d \leq 4$. Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_ho}, and if $f_0$ satisfies
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho}
	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4s/(4s + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n},
	\end{equation}
	then $\Ebb_{f_0}[1 - \phi] \leq b$.
\end{theorem}
Similarly to the first-order case, the main takeaway from Theorem~\ref{thm:laplacian_eigenmaps_testing_ho} is that $\varphi$ is a minimax optimal test over $H_0^s(P)$ when $n^{-1/2} \ll M^2 \ll n^{1/4}$ (if $2s < d$), or if $n^{-1/2} \ll M^2 \ll (2/d - 1/2)$ (if $2s \geq d$.)

However, unlike the first-order case, when $4 < d < 4s$ the minimax testing rate over $H_0^s(P)$ is still $n^{-4s/(4s + d)}$. Unfortunately, we can no longer claim that $\varphi$ is an optimal test in this regime. There do not exist any choices of $\varepsilon$ and $K$ which satisfy both~\eqref{eqn:radius_ho} and~\eqref{eqn:eigenvector_testing_ho}. As a result we can't balance testing bias and variance (defined momentarily in~\eqref{eqn:testing_biasvariance}). It follows from Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho} and the triangle inequality that, choosing $K \asymp (n)^{-2s/(2s + d)}$ as is optimal for estimation, $\varphi$ must have small Type II error whenever $\|f_0\|_P^2 \gtrsim n^{-2s/(2s + d)}$. In fact, we can derive slightly better (though still not optimal) rates.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_ho_suboptimal}
	Under the same setup as Theorem~\ref{thm:laplacian_eigenmaps_estimation_ho}, but with $d > 4$. If the Laplacian eigenmaps test $\varphi$ is computed with $K$ satisfying~\eqref{eqn:eigenvector_testing_ho}, and $\varepsilon \asymp K^{-1/d}$, and if 
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing_criticalradius_ho_suboptimal}
	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-2s/(2(s - 1) + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n},
	\end{equation}
	then $\Ebb_{f_0}[1 - \phi] \leq b$.
\end{theorem}
We do not know whether the problem here is the test itself, our some looseness in our theoretical bounds, and leave the matter for future work.

A last point: as mentioned at the beginning of this section, the estimation and testing rates we obtain in Theorems~\ref{thm:laplacian_eigenmaps_estimation_ho}-\ref{thm:laplacian_eigenmaps_testing_ho_suboptimal} are superior to those implied by the concentration of graph Laplacian eigenvectors $v_k$ around eigenfunctions $\psi_k$ of the weighted Laplace-Beltrami operator $\Delta_P$. This implies (at least) one out of the two following explanations.
\begin{enumerate}
	\item The convergence rates of $v_k \to \psi_k$ are faster than currently known,  particularly when $p$ exhibits higher-order smoothness.
	\item The approximation properties of $v_1,\ldots,v_K$ are much stronger than is implied by the concentration of each individual eigenvector $v_k$ around its continuum counterpart $\psi_k$. This is reminiscent of similar phenomena exhibited by kernel machines \textcolor{red}{(Belkin 2019)}.
\end{enumerate}
Identifying which explanation is responsible would, in our view, be a useful step towards understanding the success of graph Laplacian methods. \textcolor{red}{(TODO): Too vague and speculative. Maybe worth deleting entirely?}

@@

In this section we consider the manifold setting, where $(X_1,Y_1),\ldots,(X_n,Y_n)$ are observed according to Model~\ref{def:model_manifold}. \citet{belkin03,belkin05,niyogi2013} establish that the neighborhood graph $G_{n,r}$ can ``learn'' the manifold $\Xset$ in various senses, so long as $\Xset$ is locally linear. We build on this work by showing that when $f_0 \in H^s(P)$ and $P$ is supported on a manifold, Laplacian eigenmaps achieve the sharper minimax estimation and testing rates reviewed in Section~\ref{subsec:minimax_rates_sobolev}.

\subsection{Error rates under the manifold hypothesis}
We now show that Laplacian eigenmaps achieves the faster minimax rates under the manifold hypothesis, when the regression function $f_0 \in H^s(P;M)$, $s \leq 4$, and, in the testing problem, $m \leq 4$. This section will proceed in a similar fashion to Section~\ref{sec:higher_order_sobolev_classes}, except with the ambient dimension $d$ replaced by intrinsic dimension $m$.

\paragraph{Estimation.}
To ensure that $\wh{f}$ is an in-sample minimax rate-optimal estimator, we choose the graph radius $\varepsilon$ and number of eigenvectors $K$ as in~\ref{asmp:parameters_estimation_ho}, except with ambient dimension $d$ replaced by the intrinsic dimension $m$.
